---
title: "Data607_week7_assignment"
author: "Charls Joseph"
date: "October 11, 2018"
output:
  html_document: default
  pdf_document: default
---

## Assignment – Working with XML and JSON in R

Here is the assignmnet work statment: 

Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more
than one author. For each book, include the title, authors, and two or three other attributes that you find
interesting.
Take the information that you’ve selected about these three books, and separately create three files which
store the book’s information in HTML (using an html table), XML, and JSON formats (e.g. “books.html”,
“books.xml”, and “books.json”). To help you better understand the different file structures, I’d prefer that you
create each of these files “by hand” unless you’re already very comfortable with the file formats.
Write R code, using your packages of choice, to load the information from each of the three sources into
separate R data frames. Are the three data frames identical?
Your deliverable is the three source files and the R code. If you can, package your assignment solution up into
an .Rmd file and publish to rpubs.com. [This will also require finding a way to make your three text files
accessible from the web].



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("kableExtra")

```

### Libraries Requried
```{r includle=FALSE} 

library(XML)    
library(RCurl)  ## required for GetURL method 
library(jsonlite)
library(kableExtra)
library(knitr)


```


### Reading html file and parsing. 

htmlParse is used to parse the html content and invoked readHTMLTable to extract the html table. 

```{r html_parse_df}
html.url <- "https://raw.githubusercontent.com/charlsjoseph/CUNY-Data607/master/week7/books.html"

html.url.content= getURL(html.url)
books.html <- htmlParse(file= html.url.content )


books.html.df <- as.data.frame(readHTMLTable(books.html))
books.html.df

head(books.html.df)


# print datafrane using kabble

books.html.df %>%
  kable() %>%
  kable_styling()


```

### Reading XML file and parsing. 

xmlParse is used to parse the xml content and invoked xmlToDataFrame to convert the content into dataframe


```{r xml_parse_df}
xml.url <- "https://raw.githubusercontent.com/charlsjoseph/CUNY-Data607/master/week7/books.xml"

xml.url.content= getURL(xml.url)


xml.url.content <- xmlParse(file=xml.url.content)
xml.url.df <- xmlToDataFrame(xml.url.content)
xml.url.df

# print datafrane using kabble
xml.url.df %>%
  kable() %>%
  kable_styling()

```

### Reading JSON file and parsing. 

fromJSON is used to parse and convert into the json df


```{r json_parse_df}
json.url <- "https://raw.githubusercontent.com/charlsjoseph/CUNY-Data607/master/week7/books.json"
json.url.content= getURL(json.url)
json.url.df <- fromJSON(json.url.content)
json.url.df

# print datafrane using kabble

json.url.df %>%
  kable() %>%
  kable_styling()
```

### Conclusion

All three dataframes are almost same other than very minor difference in the column name. For e.g the column names in html df has a dot, but for xml and json  has _. Also the column name is prefixed with a NULL. 


