---
title: "Movie Ratings"
author: "Charls Joseph"
date: "September 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require("RMySQL")
require("dplyr")
require("htmlTable")
```

## Load Data frame from the mySql

### Database name :movies

### Tables : movie_ratings and movies

Make sure 'movies' and 'movie_ratings' tables present in the 'movies' database

```{r}

mydb = dbConnect(MySQL(), user='root', password='mysql@123', dbname='movies', host='localhost')
movie_ratings <- dbGetQuery(mydb, "select * from movie_ratings")
movies <- dbGetQuery(mydb, "select * from movies")
head(movie_ratings)
head(movies)

```

## inner_join(), group_by(), summarise() and sort() functions ( from dplyr package)

group_by function seprates data frame into different group based on the specified columns.
This object doesn't seem useful by itself, the real power comes when we use this with other function like summarise which does aggregation

Note: summarise() takes the data frame and apply one or more aggregation function. For eg:
avg_rating is the mean aggregation.

group_by() takes the data frame and create group based on specified columns
with() uses to order a dataframe by a specified column

```{r}

movie_ratings_join <- inner_join(movie_ratings, movies, by = c("movie_id" = "movie_id"))

#group by movie_name and movie_type
grp_movie_name_type <- group_by(movie_ratings_join, movie_name, movie_type)

#top rated movie df
top_rated_movie_name <- summarise(grp_movie_name_type, avg_rating=mean(rating, na.rm=TRUE))

#top rated 3 movies( slice operator using with() and order() function )
top_rated_movies <- head(top_rated_movie_name[with(top_rated_movie_name, order(-avg_rating)),],3)
```

## Print Top Rated movies

```{r}
htmlTable(top_rated_movies, caption = 'Top rated Movies')


```


