---
title: "MushroomDataAnalysis"
author: "Charls Joseph"
date: "August 28, 2018"
output: html_document
---


Assignment Solution for the Data607 -Week 1

Load the data from the git repository (agaricus-lepiota.data) using rcurl package. 

```{r Loading_Data}
require(RCurl)
dataLink <- getURL("https://raw.githubusercontent.com/charlsjoseph/CUNY-Data607/master/week1/607-week1-Assignment/agaricus-lepiota.data", ssl.verifypeer = FALSE)
agaricus.lepiota <- read.csv(text = dataLink)
```

Extracting the 4 columns 
Using slicing [] operator to fetch all rows and specific columns.

df[ rows, columns ]  <br>

1. Type ( edible=e, poisonous=p ) <br>
2. cap-shape  (bell=b,conical=c,convex=x,flat=f,knobbed=k,sunken=s) <br>
3. cap-surface (fibrous=f,grooves=g,scaly=y,smooth=s) <br>
4. odor (almond=a,anise=l,creosote=c,fishy=y,foul=f, musty=m,none=n,pungent=p,spicy=s) <br>


```{r Extract_Column}
edat1 <- agaricus.lepiota[, c(1:3, 6)]
head(edat1)
colnames(edat1) <- c("type", "capshape", "capsurface", "odor")
head(edat1)
```


Tranforming data using sapply function. Below code transform the abbreviation into a meaningfull name. 


```{r transform_Data}

edat1$type <- sapply(edat1$type, function(x)
{switch(as.character(x), 
        "e" = "edible", 
        "p" = "pois")
})


edat1$capshape <- sapply(edat1$capshape, function(x)
{switch(as.character(x), 
        "b" = "bell", 
        "c" = "conical", 
        "f" = "flat" , 
        "k"= "knobbed", 
        "s"= "sunken", 
        "x"="convex")
})

edat1$capsurface <- sapply(edat1$capsurface, function(x)
{switch(as.character(x), 
        "f" = "fibrous", 
        "g" = "grooves", 
        "y" = "scaly" , 
        "s"= "smooth")
})

edat1$odor <- sapply(edat1$odor, function(x)
{switch(as.character(x), 
        "a" = "almond", 
        "l" = "anise", 
        "c" = "creosote" , 
        "y"= "fishy", 
        "f" = "foul", 
        "m" = "musty" , 
        "n" = "none" , 
        "p" = "pungent", 
        "s" = "spicy")
})

```

Printing the head of transformed data. 

```{r}
head(edat1)
```
