---
title: "CleaningTidy Project Data607 "
author: "Charls Joseph"
date: "October 7, 2018"
output: html_document
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data607 Tidying and transformation Project 

In this project, I am asked to select 3 'wide' datasets and use tidyr and dplyr as needed to tidy and transform your data.

####  Library Required

Below libraries used. 
tidyr  -> provides two important functions gather() and spread()
dplyr -> provides select(), filter, group_by, summarize and mutate

```{r}

library(readr)
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(knitr)

```


## DataSet1 : Major League Baseball Team
This data set can be found on the https://www.baseball-reference.com/leagues/MLB/
and it includes every MLB team from all time, there wins, and the games played in that year.

Dataset is available in csv format to download. 

Below snippet will read the csv file and print first few samples. 

```{r read_mlb_csv}
mlbData <- data.frame(read_csv("dataset/MLB.csv"))

head(mlbData)

```

## Transform the wide format into Long format. 

Below snippet transform the dataset from wide format to long. 

```{r tranform_wide}

mlbData <- mlbData %>% gather(team, wins, 3:length(colnames(mlbData)))
head(mlbData)

```

## Tranform with dplyr funcions. 

Since the number of Games in each year varies, the points can be used for comparison purpose, So have to create a new column for win percentage(win_pct) and is calculated by total number of wins/ total no of Games.

```{r dypr_transform}

mlbData <- mlbData %>% mutate(win_pct = wins/G)
head(mlbData)

```

Print the topper and their points since 2000. 

```{r topper }

mlbData <- mlbData %>% filter(Year >= 2000)

mlbData_byyear_byteam = mlbData %>% group_by(Year, team)

mlbData_byyear = mlbData %>% group_by(Year) %>% summarise( max_wins = max(wins, na.rm = TRUE))  %>% arrange(-Year)

topper <- inner_join(mlbData_byyear, mlbData_byyear_byteam,   c("Year" = "Year", "max_wins" = "wins")  ) %>% select(Year, max_wins , team)
kable(topper)

```




## DataSet2 : Sample dataset of babies height from year 2011 to year 2015.

Below snippet will read the csv file and print first few samples. 


```{r read_csv}
babiesHeightData <- data.frame(read_csv("dataset/babies_heights.csv"))

head(babiesHeightData)

```

## Format the column name to year 

```{r }

colnames(babiesHeightData)[3:7] <- c("2011", "2012" , "2013" , "2014" , "2015")
colnames(babiesHeightData)


```


## Transform from wide to long dataset 

Using tidyr gather() method, transform the wide format to long. Year column value becomes row data value. 

```{r}
babiesHeightData <- babiesHeightData %>% gather(year, height, 3:7)

head(babiesHeightData)

```

## Average heights by year and by age. 

using dplyr( using pipe), calculate the average heights by year and by age. 


```{r}

heights_by_gender_by_year = babiesHeightData %>% group_by(year,gender) %>% summarise(mean_height = mean(height, na.rm = TRUE))

heights_by_gender_by_year

```

## Plot average height for each year( for Male and Female)

Below R snippet plot a bar chart showing year on the X-axis and average heights on the Y-axis 

```{r}

ggplot(heights_by_gender_by_year, aes(year, mean_height)) + 
  geom_bar(aes(fill = gender), position = "dodge", stat = "identity") + 
  ylab("Average Height")



```