---
title: "CleaningTidy Project Data607 "
author: "Charls Joseph"
date: "October 7, 2018"
output: html_document
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data607 Tidying and transformation Project 

In this project, I am asked to select 3 'wide' datasets and use tidyr and dplyr as needed to tidy and transform your data.

####  Library Required

Below libraries used. 
tidyr  -> provides two important functions gather() and spread()
dplyr -> provides select(), filter, group_by, summarize and mutate

```{r}

library(readr)
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(knitr)

```


## DataSet 1 : Major League Baseball Team
This data set can be found on the https://www.baseball-reference.com/leagues/MLB/
and it includes every MLB team from all time, there wins, and the games played in that year.

Dataset is available in csv format to download. 

Below snippet will read the csv file and print first few samples. 

```{r read_mlb_csv}
mlbData <- data.frame(read_csv("https://raw.githubusercontent.com/charlsjoseph/CUNY-Data607/master/week5/project/dataset/MLB.csv"))

head(mlbData)

```

## Transform the wide format into Long format. 

Below snippet transform the dataset from wide format to long. 

```{r tranform_wide}

mlbData <- mlbData %>% gather(team, wins, 3:length(colnames(mlbData)))
head(mlbData)

```

## Tranform with dplyr funcions. 

Since the number of Games in each year varies, the points can be used for comparison purpose, So have to create a new column for win percentage(win_pct) and is calculated by total number of wins/ total no of Games.

```{r dypr_transform}

mlbData <- mlbData %>% mutate(win_pct = wins/G)
head(mlbData)

```

Print the topper and their points since 2000. 
used join function to join two dataset to get the top team name and its max score 

```{r topper }

mlbData <- mlbData %>% filter(Year >= 2000)

mlbData_byyear_byteam = mlbData %>% group_by(Year, team)

mlbData_byyear = mlbData %>% group_by(Year) %>% summarise( max_wins = max(wins, na.rm = TRUE))  %>% arrange(-Year)

topper <- inner_join(mlbData_byyear, mlbData_byyear_byteam,   c("Year" = "Year", "max_wins" = "wins")  ) %>% select(Year, max_wins , team)
kable(topper)

```




## DataSet 2 : Sample dataset of babies height from year 2011 to year 2015.

Below snippet will read the csv file and print first few samples. 


```{r read_csv_babies_hieghts}
babiesHeightData <- data.frame(read_csv("https://raw.githubusercontent.com/charlsjoseph/CUNY-Data607/master/week5/project/dataset/babies_heights.csv"))

head(babiesHeightData)

```

## Format the column name to year 

```{r }

colnames(babiesHeightData)[3:7] <- c("2011", "2012" , "2013" , "2014" , "2015")
colnames(babiesHeightData)


```


## Transform from wide to long dataset 

Using tidyr gather() method, transform the wide format to long. Year column value becomes row data value. 

```{r}
babiesHeightData <- babiesHeightData %>% gather(year, height, 3:7)

head(babiesHeightData)

```

## Average heights by year and by age. 

using dplyr( using pipe), calculate the average heights by year and by age. 


```{r}

heights_by_gender_by_year = babiesHeightData %>% group_by(year,gender) %>% summarise(mean_height = mean(height, na.rm = TRUE))

heights_by_gender_by_year

```

## Plot average height for each year( for Male and Female)

Below R snippet plot a bar chart showing year on the X-axis and average heights on the Y-axis 

```{r}

ggplot(heights_by_gender_by_year, aes(year, mean_height)) + 
  geom_bar(aes(fill = gender), position = "dodge", stat = "identity") + 
  ylab("Average Height")



```


## DataSet 3 : High School and Beyond Data Set

This dataset is found in https://rdrr.io/cran/blorr/man/hsb2.html
Below snippet will read the csv file and print first few samples. 


id -
id of the student

female -
gender of the student

race -
ethnic background of the student

ses -
socio-economic status of the student

schtyp -
school type

prog - 
program type

read
scores from test of reading

write - 
scores from test of writing

math - 
scores from test of math

science - 
scores from test of science

socst - 
scores from test of social studies




```{r read_csv_hsb2}
hsb2Data <- data.frame(read_csv("https://raw.githubusercontent.com/charlsjoseph/CUNY-Data607/master/week5/project/dataset/hsb2.csv"))

head(hsb2Data)

```


## Transform from wide to long dataset 

Using tidyr gather() method, transform the wide format to long. Year column value becomes row data value. 

```{r}
hsb2Data <- hsb2Data %>% gather(ability, score, read:socst)

head(hsb2Data)

```

## Tranform using dplyr 

create a bar plot with socio-economic status of the student in X axis and average score of different abilities 
to understand if socio-economic status has any relation with their ability. 

For this we can make out that students with low socio-economic status having the score < 50 
and students with high socio-economic status has comparatively high average score.


```{r }
hsb2Data$ses.f = factor(hsb2Data$ses, labels=c("low", "middle", "high")) 

meanScore_byses.f_byability = hsb2Data %>% group_by(ses.f, ability) %>% summarise(mean_score = mean(score, na.rm = TRUE))

head(meanScore_byses.f_byability)

ggplot(meanScore_byses.f_byability, aes(ses.f, mean_score)) + 
  geom_bar(aes(fill = ability), position = "dodge", stat = "identity") + 
  ylab("Average Score")



```

