---
title: "R Regex and String Manipulation"
author: "Charls Joseph"
date: "September 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
```

## R Markdown


### Question 3. 
Copy the introductory example. The vector name stores the extracted names.
R> name
[1] "Moe Szyslak" "Burns, C. Montgomery" "Rev. Timothy Lovejoy"
[4] "Ned Flanders" "Simpson, Homer" "Dr. Julius Hibbert"
(a) Use the tools of this chapter to rearrange the vector so that all elements conform to
the standard first_name last_name.

```{r}
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555
-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson,
Homer5553642Dr. Julius Hibbert"

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
step1 <- sub(" [A-z]{1}\\. "," ",name)  # remove initials
step2 <- sub("(\\w+),?\\s(\\w+)","\\2 \\1", step1) # switch last,first to first last 
step3 <- sub("[A-z]{2,}\\. ","",step2)  # remove titles
standard_name <- gsub('^[[:punct:]]|[[:punct:]]$', '', step3) # trim any punctuations if any. 

standard_name

```

(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and
Dr.).

```{r}
title <- str_detect(name,"[A-z]{2,}\\. ") #logical vector to show the title.
title

```



(c) Construct a logical vector indicating whether a character has a second name.

```{r}

second_name <- str_detect(name," [A-z]{1}\\. ") #logical vector to show the second name. 
second_name

nameDF <- data.frame(standard_name,title, second_name)
names(nameDF) <- c("Name" , "Title", "Second_Name")
nameDF
```



###Question 4. 
Describe the types of strings that conform to the following regular expressions and
construct an example that is matched by the regular expression.
(a) [0-9]+\\$
Regex to extract the digit sufixed with $ 


```{r }
strA <- "The price of my watch is 67$"

dollarValues <- str_extract_all(strA , "[0-9]+\\$")
dollarValues
```


(b) \\b[a-z]{1,4}\\b

This reges expression extract a word matches 1 to 4 character lower case letters. 
 

```{r}
strB <-  "Lord says, My grace is sufficient for you, for my power is made perfect in weakness. Therefore I will boast all the more gladly about my weaknesses, so that Christ power may rest on me."

words <- str_extract_all(strB , "\\b[a-z]{1,4}\\b")
words

```



(c) .*?\\.txt$

This regex extract the any character pattern ending with .txt. This can be used to fetch the filenames. 

```{r}

filenames <- "Automatic_data_collection.txt"
files <- str_extract_all(filenames , ".*?\\.txt$")
files

```

(d) \\d{2}/\\d{2}/\\d{4}

This Regex extract the date pattern dd/mm/yyyy. 

01/01/1982

```{r}

numericalue <- "56/56 20/11/2017  "
date_pattern <- str_extract_all(numericalue , "\\d{2}/\\d{2}/\\d{4}")
date_pattern

```



(e) <(.+?)>.+?</\\1>

This Regex matches the open and close tag and the value inside the tag. 

```{r}

htmltag <- "5635 <name>charls joseph</name>38787"
tags <- str_extract_all(htmltag , "<(.+?)>.+?</\\1>")
tags
  
  
```

### Question 9:
The following code hides a secret message. Crack it with R and regular expressions.
Hint: Some of the characters are more revealing than others! The code snippet is also
available in the materials at www.r-datacollection.com.

clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr

```{r}
message <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

secret_message <- str_extract_all(message , "[[:upper:][:punct:]]")
secret_message

```


