---
title: "Project1 - Tournament results"
author: "Charls Joseph"
date: "September 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(stringr)
```

## Step-1 : Extract the row data using Regex expression.


Line1 is extracted using Regex and used couple of replace code to remove unwanted spaces and characters. 
Use lapply to apply a series of replace functions to remove unwanted characters and extra space and end result will be delimited with "|"

```{r extract_first_lines}


mystring <- read_file("https://raw.githubusercontent.com/charlsjoseph/CUNY-Data607/master/week3/tournamentinfo.txt")


#extract the 1st Line using regex 
lines1 <- str_extract_all(mystring, "\\w+.+[|]\\d+[.]\\d.+\\|")

firstlines <- lapply(lines1, function (x){
  x <- str_replace_all(x, "[|][A-Z]\\s{1,}", "|")
  x <- str_replace_all(x, "\\s{2,}" , "")
  x<- str_replace_all(x, "[|]{2}" , "|0|")
  str_replace_all(x, "[|]{2}" , "|0|")
})

head(firstlines)
```

## Step-2 : Extract the names, points and calculate the average Points from first Line. 


Format the firstLines and extract the data into names, points and averagePoints vector variables.
notes: used str_split() function to split the line with the delimiter "|"

For e.g 
1 | GARY HUA|6.0|39|21|18|14|7|12|4|
Extract following into names, points, averagePoints vector variables

names ="GARY HUA"
points = 6.0
averagePoints = mean(21,18,14,7,12,4)

```{r format_names_point_avg}
names <- c()
id <- c()
points <- c()
round1 <- c()
round2 <- c()
round3 <- c()
round4 <- c()
round5 <- c()
round6 <- c()
round7 <- c()
for (row in unlist(firstlines)) {
  ar<- unlist(str_split(row, fixed("|")))
  id <- c(id, as.integer(ar[1]))
  names <- c(names, ar[2])
  points <- c(points, ar[3])
  print(row)
  round1 <- c(round1, as.integer(ar[4]))
  round2 <- c(round2, as.integer(ar[5]))
  round3 <- c(round3, as.integer(ar[6]))
  round4 <- c(round4, as.integer(ar[7]))
  round5 <- c(round5, as.integer(ar[8]))
  round6 <- c(round6, as.integer(ar[9]))
  round7 <- c(round7, as.integer(ar[10]))
}
length(round7)

```

## Step-3 : Extract the state and the Pre-rating data from second line 

### 1. Extract state data and store in statesData vectors  

```{r extract_state_data} 

statesData <- str_extract_all(mystring, ".+[A-Z]{2}\\s{1}[|]")
states = c()
for (state in unlist(statesData)) {
  state <- str_replace_all(state, "\\s+" , "")
  state <- str_replace_all(state, "[|]" , "")
  states <- c(states, state)
}

states
length(states)

```


### 2. Extract Pre-rating data and store into prepostData   

```{r extract_preRateData}
preRating = c()
prepostData <- str_extract_all(mystring, "[0-9,A-Z]+\\s*->\\s?[0-9,A-Z]+")
for (prepost in unlist(prepostData)) {
  prepost <- str_replace_all(prepost, "P?[\\d+]?.?->" , "|")
  ar<- unlist(str_split(prepost, fixed("|")))
  preRating <- c(preRating,as.integer(ar[1]))
}
preRating
```

## Step-4 Create a dataframe with following columns. 

id=id
names
points
oponent1
oponent2
oponent3
oponent4
oponent5
oponent6
oponent7
state 
preRating

```{r create_df}
players.data <- data.frame(id=id,names=names,points=points, oponent1=round1,
                           oponent2=round2,
                           oponent3=round3,
                           oponent4=round4,
                           oponent5=round5,
                           oponent6=round6,
                           oponent7=round7,
                          state=states ,preRating=preRating  )

head(players.data)

```

## Final Step: Calculate the average Prerating of Opponents

Created a fn called fetch_PreRating_oponent to fetch the prerating for the opponent

#### fn: fetch_PreRating_oponent

Input Parameter: Opponent Id 

OutPut Parameter : PreRating Value


use Apply function to generate a new calculated column called "avgPreRating"


```{r final_step}

fetch_PreRating_oponent <- function(oponent) {
  
  df= players.data[players.data[1] == oponent ]
  c <- as.integer(df[12])
  if (is.na(c)) {
        c <- 0
  }
  return(c)
}

players.data$avgPreRating <- apply(players.data, MARGIN=1, function(x){
trunc(mean(c(fetch_PreRating_oponent(as.integer(x[4])), fetch_PreRating_oponent(as.integer(x[5])), fetch_PreRating_oponent(as.integer(x[6])), fetch_PreRating_oponent(as.integer(x[7])), fetch_PreRating_oponent(as.integer(x[8])), fetch_PreRating_oponent(as.integer(x[9])), fetch_PreRating_oponent(as.integer(x[10])))))

})

finalResult <- players.data[,c("names", "state" , "points", "avgPreRating")]
write.csv(finalResult, "finalResults.csv")
finalResult
```

