---
title: "Tidying and transformation"
author: "Charls Joseph"
date: "September 30, 2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

This assignment is to read the arrival delays for two airlines across five destination from a csv file. 
The data needs to be transformed using tidyr methods spread() and gather() in order to analyse the delay rate between two airlines. 

## Library Required

```{r}

library(readr)
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)

```

## Read csv and fill the missing column names

```{r read_csv}
flightsData <- data.frame(read_csv("https://raw.githubusercontent.com/charlsjoseph/CUNY-Data607/master/week5/FlightData.csv"))

head(flightsData)

```

Added flightname and the status into column. 
flightname is missing on the alternate row. 

```{r missing_data}

flightsData[1] <- rep(flightsData[which(!is.na(flightsData[1])), 1], each = 2)

colnames(flightsData)[1:2] <- c("FlightName" , "status")
head(flightsData)

```

## Using tydyr metods Gather() and Spread() which enlarges the data. 

Gather() puts column values into row data.  desitination, count are the new columns 

```{r tidy_Gather}
flightsData1 <- flightsData %>% gather(desitination, count, 3:7)

head(flightsData1)

```

Spread() puts row value into column. values on status is the new columns

```{r tidy_Spread}
flightsData2 <- flightsData1 %>% spread(status, count)

colnames(flightsData2) <- str_replace(colnames(flightsData2), " ", "_")


head(flightsData2)

```

## Calculate delay Rate for a flight 

delay Rate is calculated by the  delayed / (delayed + on_time
```{r}
flightsData3 <- flightsData2 %>% mutate(delay_rate = delayed / (delayed + on_time))
head(flightsData3)

```

## Plot the chart 
Plot the chart with X axis as flights and y as Delay Rate. From the chart, it is evident that the "Am WEST" is having higher Delay Rate for all destination locations. 

```{r}


ggplot(flightsData3, aes(desitination, delay_rate)) + 
  geom_bar(aes(fill = FlightName), position = "dodge", stat = "identity") + 
  ylab("Delay Rate")

```
